---
title: "united_states_budget"
author: "Benjamin Panny"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's look at [Table 4.1—Outlays by Agency: 1962–2029](https://www.whitehouse.gov/omb/budget/historical-tables/)

```{r, message = F}
library(httr)
library(readxl)
library(tidyverse)

# URL of the Excel file
url <- "https://www.whitehouse.gov/wp-content/uploads/2024/03/hist04z1_fy2025.xlsx"

# Download file into memory
response <- GET(url)

# Create a temporary file
temp_file <- tempfile(fileext = ".xlsx")

# Write the content of the response to the temporary file
writeBin(content(response, "raw"), temp_file)

# Read the Excel file from the temporary file
file <- read_excel(temp_file)

# (Optional) Clean up by deleting the temporary file after use
unlink(temp_file)

file %>% head
```


```{r}
library(tidyverse)
file[1,] %>% t() %>% unique
```

Row 1 only mentions that values are given in millions of dollars, so it can be dropped. Also, "Department or Other Unit" is the true column label for the rest of what is in column 1, while the rest of row 2 is the year. I eventually want to plot by year and by department. So let's transform the dataframe to have a column for Department, a column for Year, and a column for Outlay

```{r}
data <- file[-1,] %>% 
  rename(tmp_colname = `Table 4.1 - OUTLAYS BY AGENCY:  1962 - 2029`) %>% 
  mutate(tmp_colname = ifelse(tmp_colname == "Department or other unit", "Year", tmp_colname)) %>% 
  t() %>% 
  as_tibble()

clean_colnames <- unlist(data[1,])
data <- data[-1,]
colnames(data) <- clean_colnames
dep_of_interest <- clean_colnames[1:34]

data_long <- data %>% 
  pivot_longer(cols = -Year,
               names_to = "Department or Other Unit",
               values_to = "Outlay") %>% 
  # TQ stands for "Transition Quarter" and dirties my Year column type!
  filter(Year != "TQ") %>% 
  mutate(across(everything(), ~na_if(.,"..........")),
         Year = str_replace(Year, " estimate", "") %>% as.integer(),
         Outlay = as.numeric(Outlay)) %>% 
  replace_na(list("Outlay" = 0)) %>% 
  # I don't want to deal with negative outlays, so I'm removing them
  mutate(Outlay = ifelse(Outlay < 0, NA, Outlay)) %>% 
  filter(!is.na(Outlay), `Department or Other Unit` %in% dep_of_interest)

data_long %>% 
  filter(str_detect(`Department or Other Unit`, 'Department')) %>% 
  ggplot(aes(x = Year, y = Outlay, fill = `Department or Other Unit`)) +
  geom_area(stat = "identity", position = "fill") +
  # for totals instead of proportions
  # geom_area()
  theme(legend.position = "top")
```

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(stringr)

# Assuming data_long is already loaded and prepared
data_long %>%
  filter(str_detect(`Department or Other Unit`, 'Department')) %>%
  ggplot(aes(x = Year, y = Outlay, fill = `Department or Other Unit`, text = `Department or Other Unit`)) +
  geom_area() +
  theme(legend.position = "none") -> p

# Convert ggplot to plotly for interactivity
interactive_plot <- ggplotly(p, tooltip = "text")

# Display the plot
interactive_plot
```

```{r}
# # Library
# library(streamgraph)
# 
# # Stream graph with a legend
# pp <- streamgraph(data_long %>%
#                     filter(str_detect(`Department or Other Unit`, 'Department')) %>% 
#                     as.data.frame() %>% 
#                     rename("key" = `Department or Other Unit`,
#                            "value" = "Outlay",
#                            "date" = "Year"), 
#                  height="300px", width="1000px") %>%
#   sg_legend(show=TRUE, label="names: ")
# 
# 
# # save the widget
# library(htmlwidgets)
# saveWidget(pp, file=paste0( getwd(), "/streamgraphDropdown.html"))
```

